---
title: "Highlands Parking"
date: "`r Sys.time()`"
output: html_notebook
author: "Rick Gilmore"
---


```{r}
library(tidyverse)
```

# Purpose

To visualize data about garage, on-street, and lawn parking on football weekends.

# Import & clean data

```{r import-data}
garage <- read.csv("csv/spaces-avail.csv")
lawn <- read.csv("csv/lawn-parking.csv")
street <- read.csv("csv/street-parking.csv")
```

The games are in separate columns. It's easier to visualize patterns if these are converted to one column with different labels.

```{r munge-into-column}
garage %>% 
  gather(., game, spaces, game1:game7) ->
  garage_clean
garage_clean
```

Let's do the same for the other datasets.

```{r clean-lawn}
lawn %>% 
  gather(., game, cars, game1:game7) ->
  lawn_clean
```

```{r clean-street}
street %>% 
  gather(., game, cars, game1:game7) ->
  street_clean
street_clean
```

# Plots

```{r plot-lawn}
lawn_clean %>%
  ggplot(.) +
  aes(x=time, y=cars, fill=neighborhood) +
  geom_col() +
  facet_grid(neighborhood ~ game) +
  ggtitle("Cars on lawns by neighborhood")
```

```{r plot-street}
street_clean %>%
  ggplot(.) +
  aes(x=time, y=cars, fill=neighborhood) +
  geom_col() +
  facet_grid(neighborhood ~ game) +
  ggtitle("Cars on street by neighborhood")
```

```{r plot-garage}
garage_clean %>%
  ggplot(.) +
  aes(x=time, y=spaces, fill=garage) +
  geom_col() +
  facet_grid(garage ~ game) +
  ggtitle("Available spaces by garage")
```

```{r nuisance-parking}
street_totals <- street_clean %>%
  filter(time %in% c("am", "pm")) %>%
  select(game, time, cars) %>%
  group_by(game, time) %>%
  summarize(n_cars = sum(cars))

lawn_totals <- lawn_clean %>%
  select(game, time, cars) %>%
  group_by(game, time) %>%
  summarize(n_cars = sum(cars))

nuisance_cars <- street_totals$n_cars + lawn_totals$n_cars
nuisance <- data_frame(game = lawn_totals$game, time = lawn_totals$time, cars = nuisance_cars)

nuisance %>% 
  ggplot(.) +
  aes(x=time, y=cars, fill=time) +
  geom_col() +
  facet_grid(. ~ game) +
  ggtitle("Cars on street or lawn")
```

# Observations

- On busy game weekends, there are 500-2,000+ cars parked on Borough streets and lawns, most of these in the Highlands and Holmes-Foster neighborhoods.

- There is insufficient capacity in the Borough garages to accommodate this volume.
 